
\documentclass[../thesis.tex]{subfiles}

\begin{document}

\chapter*{Abstract}
\textbf{Background:}
Increasingly many programmers become unable to work
due to developing pain from using computers
too much over a long period of time.
Thanks to recent developments, voice control has become a very viable option
for people with a variety of conditions.
At the same time, tools that easily enable
analysis of programs in many different languages,
such as Language Servers and TreeSitter, are becoming readily available.
\newline
\textbf{Aim:} 
This thesis aims to investigate how different tools
can be integrated with general purpose voice control systems
in order to increase the efficiency with programmers using voice control.
\newline
\textbf{Method:} A functional prototype of an extension to Talon was developed and evaluated through
a qualitative user study of 6 participants.
\newline
\textbf{Results:} The results showed that participants
had a very positive impression of the prototype.
The prototype itself also shows that TreeSitter is an excellent tool
for the kind of syntactic analysis used in this project.
\newline
\textbf{Conclusion:} 
The techniques presented in this thesis
can indeed improve the efficiency of programmers using voice controls.

% Software developers spend a lot of time working on a computer.
% This makes them particular susceptible to Repetitive Strain Injuries (RSI), which can be detrimental to one's career.
% Recently there have been great advancements in hands-free input systems that allow users to control a computer using their voice and/or an eye-tracker.
% These systems can be customized using one or more scripting languages, which allows users to acquire much higher proficiency than what was previously thought possible for someone without full use of their hands.
% There is however a steep learning curve, and a lot of work that goes into iteratively creating voice commands that allows for a fluent workflow.
% The goal of this thesis is to uncover techniques to automate this process, and increase the productivity of programmers coding with their voice.

\end{document}
