
\appendix
\begin{appendices}
\chapter{Interview Questions}
% The contents...
The interview consisted of two rounds of questions that was asked
before and after the user test.

\section{Questions Round 1}
\begin{enumerate}[label*=Q1.\arabic*]
    \item How long have you been programming?
    \item How long have you been using Talon or other Voice Coding systems?
    \item How much do you use Talon compared to keyboard and mouse in your day to day work?
    \item Which speech engine are you using?
    \item What command set are you using (is it based on knausj)?
    \item Are you familiar with Elm or any similar languages such as Haskell or SML?
    \item Do you have experience using Vim?
    \item What are some issues or challenges you experience when programming with Talon?
\end{enumerate}

\newpage
\section{User Test}

After having installed the system, the user was asked do try the following
navigational and editing commands:
\begin{minted}[frame=lines]{text}
go fun update
go type update
go to sandbox

go to text
go to next
go to last

go string
go number
go import

select type person
select fun view

delete type message
save it
delete fun main
\end{minted}
They were then asked to try the following dictation commands:
\begin{minted}[frame=lines]{text}
word person <pause> of type person 
--> person : Person
word update <pause> of type message to model to model 
--> update : Msg -> Model -> Model
\end{minted}
\section{Questions Round 2}
\begin{enumerate}[label*=Q2.\arabic*]
    \item In general, what is your impression of what you have seen during this session?
    \item Were there any feature you thought was particularly useful?
    \item Were there any behavior you found to be strange or unintuitive?
    \item Are there any features that are not implemented that you think could improve the system?
    \item Do you feel like the system addresses some of the challenges you have been experiencing with vocal programming?
    \item Based on what you have seen so far, would you be interested in using a finished version of a system like this?
\end{enumerate}


\end{appendices}
